---
title: "Housing Data - Learning Methods"
author: "Ryan Monks"
date: "20 January 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# House Prices: Advanced Regression Techniques

This is a dataset found on Kaggle often used to learn new methods and techniques. The data is split into a test and a training set. This Ames dataset was compiled for use in data science education.


## Initial Overview of dataset
There are 80 predictive variables in total and 1460 rows, only factors and integers. There are no numeric variables with decimal places.


```{r}
train <- read.csv("C:/Users/Ryan/Desktop/R folder/House Prices/train.csv")
test <- read.csv("C:/Users/Ryan/Desktop/R folder/House Prices/test.csv")
library(ggplot2)
str(train)
```
There are too many columns to review plots inidividually.

## Missing Data

Understanding how many missing values is essential for prediction and for understanding the columns. Let's have a look at the number of missing values.

```{r}
colSums(sapply(train,is.na))
```
There are a significant number of missing values for:

*PoolQC
*Fence
*MiscFeature
*GarageQual
*GarageCond
*Alley
*LotFrontage
*FireplaceQu

Summary can show us the number of NAS
```{r}
summary(train$PoolQC)
summary(train$Fence)
```

However, one Kaggle guide has shown a very good method of showing NAs:
https://www.kaggle.com/notaapple/detailed-exploratory-data-analysis-using-r

```{r}
plot_Missing <- function(data_in, title = NULL){
  temp_df <- as.data.frame(ifelse(is.na(data_in), 0, 1))
  temp_df <- temp_df[,order(colSums(temp_df))]
  data_temp <- expand.grid(list(x = 1:nrow(temp_df), y = colnames(temp_df)))
  data_temp$m <- as.vector(as.matrix(temp_df))
  data_temp <- data.frame(x = unlist(data_temp$x), y = unlist(data_temp$y), m = unlist(data_temp$m))
  ggplot(data_temp) + geom_tile(aes(x=x, y=y, fill=factor(m))) + scale_fill_manual(values=c("white", "black"), name="Missing\n(0=Yes, 1=No)") + theme_light() + ylab("") + xlab("") + ggtitle(title)
}
plot_Missing(train[,colSums(is.na(train)) > 0])
```

For some reason, the with= parameter is not needed anymore that is in the link. This code is creating a function plot_Missing that has parameter of the dataset, in our case those that are not NA.

Observations:
* A large proportion of houses miss alleys, pools, fences, fireplaces and elevators. This is consistent with what one may expect.
* The fireplace and fireplacequ variables could be combined into one.

## Cleaning Data
One good idea is to convert some character variables to numeric variables to allow for quicker assignment for the model. Initially I started by assigning categorical levels with a high frequency in one bucket into binary indicators.

```{r}
train$paved[train$Street == "Pave"] <-1 
train$paved[train$Street != "Pave"] <-0
table(train$Street)

table(train$LotShape)
train$regshape[train$LotShape == "Reg"] <-1
train$regshape[train$LotShape != "Reg"] <-0
table(train$LotShape)

train$regshape[train$LandContour == "Reg"] <-1
train$regshape[train$LandContour != "Reg"] <-0
table(train$LandContour)
```
Upon looking at Garagetype as a variable to consider, it became apparent that some are more indicative of value than others as shown below. Therefore it's important to use this information.

```{r echo=FALSE}
library(knitr)
library(dplyr)
kable(summarize(group_by(train,GarageType),mean(SalePrice,na.rm=T)))
```

The above table shows that BuiltIn and Attached garages correlate to a significantly higher sale value than the other components. Therefore it's appropriate to combine these into one.

```{r}

train$GarageType1[train$GarageType %in% c("Attchd","BuiltIn")] <- 1
train$GarageType1[!train$GarageType %in% c("Attchd","BuiltIn")] <- 0
```

Repeating this for GarageFinish and GarageQual, noting that garage quality had a significant deviation between each various level.

```{r}
train$garagequal[train$GarageQual =="Ex"] <-1
train$garagequal[train$GarageQual =="Fa"] <-2
train$garagequal[train$GarageQual =="Gd"] <-3
train$garagequal[train$GarageQual =="TA"] <-4
train$garagequal[train$GarageQual == "Po" | is.na(train$GarageQual)] <- 5

train$garagefinish[train$GarageFinish %in% c("Fin", "RFn")] <- 1
train$garagefinish[!train$GarageFinish %in% c("Fin","RFn")] <- 0
```




## Correlations
```{r}
library(corrplot)
correlations<-cor(train[,c("SalePrice","OverallQual","OverallCond")])
corrplot(correlations, method="circle", type="lower",  sig.level = 0.01, insig = "blank")
```


## Plotting data
TBC

## Estimation methods


















## R Tools learnt along the way:

###Subsetting
This is the action of subsetting vectors for the desired entries. There are apparantely 5 main ways of doing so:

The method used was using the letters vector

```{r}
letters
```

#### Method 1 & 2 By position and by exclusion
The simplest is to identify by the relative position in the vector
```{r}
letters[c(1,13)]
```
Note we have to use a column vector as simply writing 1, 13 doesn't make any sense.

```{r}
letters[-c(1,13)]
```
Negative numbers mean extract all but those figures. Note that you cannot mix positive and negative.

Pros:
Easy 

Cons:
Must know position of the entry wanted

#### Method 3 - By Name

```{r}
ip <- c("Ryan"=1,"Monks"=3,"Edward"=5)
ip[c("Ryan","Edward")]
```

#### Method 4 - By logicals
Like the names example, this one is used if the position isn't known but we sort of know what we are looking for.

This method uses the [ ] operator and returns any subset of the vector containing the elements corresponding to TRUE values in our logical 'indexer'.

```{r}
x = c("a", "b", "c", "b")
x[c(TRUE, FALSE, TRUE, FALSE)]
```

This shows that it only pulls out elements 1 and 3 because they are considered true by the indexer.

```{r}
x = c("a", "b", "c", "b")
x[x=="b"]
```
Or we could find all the positions of the vector

```{r}
x = c("a", "b", "c", "b")
x[(1:length(x))[x=="b"]] 
```

What this does is produces a vector of true false for x=="b" and then that is applied to the vector (1:length(x)). This produces the same results but in a different method, albeit.

Here's another more technical example:
```{r}
x = rnorm(100, mean=4, sd=1)
x[x<1 & x>6] 
```

We can also use logicals to include or exclude values, using !. This inverts the logical vector.
```{r}
x = c("ryan", "edward", "monks")
x[!(x=="edward")] 
```

#### Method 5 - The degenerate case
This is simply doing x[ ] which returns the vector itself. This is useful if we wish to re assign the values of every entry of the vector.

URL found here: https://www.stat.berkeley.edu/~nolan/stat133/Fall05/notes/RSubset.html
